% =============================================================================
% Abstract
% =============================================================================

\begin{abstract}
We ask what structure is forced into existence by a single admissibility
criterion---\textbf{Completed Witnessability (A0$^*$)}---which requires
that every meaningful distinction be backed by a finite, replayable,
self-accounting witness process whose own validity is endogenously
checkable.

A0$^*$ is not a postulated axiom: it is \emph{derived} from absolute
nothingness ($\noth$) through a chain of five necessity theorems
(N0--N4), making it the first theorem of the framework.
This closure move forces an \emph{ordered
witness algebra} $\WitAlg$ carrying finite description, replay,
termination, separation, disturbance accounting, endogenous validity,
compositional closure, and quotient invariance.  Noncommutativity is
internal from the start; commutativity is a derived sector property.

Starting from $\noth$, we give a 34-step derivation
(Steps~0--33, nine phases A--I) that is forced up to computable
isomorphism.  Self-application of A0$^*$ to its own products yields
three primitives---tests, truth, and time---together with a carrier,
an ordered witness ledger, a truth quotient with observable-open topology,
gauge invariance, a witness-path metric (with symmetric separative
distance and directed time quasi-metric), Bellman-optimal test selection
governed by the Coupling Law (forced as the unique prior-free policy),
self-hosting closure, a universal execution law with consciousness
projector, and a K\"{a}hler geometric structure on the continuum limit
derived from a single witness generator.  A C$^*$-algebraic completion
yields a GNS Hilbert-space representation with the Born rule as the
unique probability assignment on the event algebra.

The full $\Pit$-fixed closure whole $U = \mathrm{Fix}(\Pit)$ is
atemporal: the universe is static globally, dynamic only as local
self-resolution of witness-defect.  Consciousness is derived as
self-remodelling witness-closure: a subsystem whose witness policy is
endogenously updated through an internal self-valuation channel.

The conductance exponent ($w = C/d^2$) is forced by scale covariance of
witness energy.  The Nijenhuis tensor vanishes ($N_J = 0$) by
replay-completeness, not by an external smoothness assumption.  The
spatial dimension $n = 3$ follows from matching the forced conductance
law to radial flux scaling; together with the irreversible ledger time
coordinate, the operational spacetime is $3{+}1$.  The K\"{a}hler split
law $X = J\nabla E - \nabla D$ decomposes dynamics into reversible
(quantum), connection (gauge), metric-modulation (gravity), and
dissipative (thermodynamic) sectors, recovering the Schr\"{o}dinger
equation, Yang--Mills equations, and Einstein field equations as sector
readouts.

The gauge group $SU(3) \times SU(2) \times U(1)$ is derived without
empirical input: rank~$1$ from the K\"{a}hler complex structure, rank~$2$
from the $3{+}1$ spin structure, rank~$3$ from anomaly cancellation.

The quantitative seed $\delta^*$ is the unique least positive
self-retaining non-gauge closure-defect minimizing the action functional.
Every dimensionless observable is forced as a spectral, holonomy, or
normal-form invariant of the linearized refinement operator $L^*$ at
$\delta^*$.  Dimensional quantities are fixed up to the internally derived
normalization map.

We complete the source code operationally
(\cref{sec:manifestability}) by introducing the
\emph{refinement threshold} $\chi(W)$ of an unresolved
indistinguishability class and the \emph{refinement kernel}
$\mathcal{K}$, making the universe not only structurally and
quantitatively forced, but also operationally governed by the
least-action, value-weighted refinement of hidden distinctions.
The Bellman/Hamilton--Jacobi value equation
$\Psi(W) = \sup[V - A + \sum \Psi(W_i)]$ is the exact operational law.
As a consequence, the true computational state of any NP verifier is
its residual future-equivalence quotient, whose polynomial size is
forced by A0$^*$ spectral curvature, yielding P $=$ NP as an
inevitable structural consequence.

No modeling commitment or empirical input survives: closed universe,
commutativity, prior-free selection, conductance exponent,
efficiency-landscape smoothness, and the gauge group are all derived from
A0$^*$---which is itself derived from $\noth$.  The framework has no
declared axioms.  The derivation ($\sim$120 pages) is accompanied by a
three-point self-audit, Z3 verification on finite models, 140 Lean~4
files (zero sorry, one axiom A0$^*$),
and three falsifiable predictions.
\end{abstract}
