% =============================================================================
% Section: Qualitative and Quantitative Forcing from Nothingness
% =============================================================================

\section{Qualitative and Quantitative Forcing}
\label{sec:forcing}

Before proceeding to geometry and physics, we prove the two master theorems
that bridge nothingness to structured reality.  From $\noth$, there are no
external labels, names, coordinates, units, or scales.  This forces all
qualitative structure to be quotient-invariant and all quantitative
structure to be internal-comparison-invariant.

% ---------------------------------------------------------------------------
\subsection{Qualitative Forcing}
\label{sec:forcing:qualitative}

\begin{theorem}[Q1 --- Qualitative Forcing]
\label{thm:qualitative-forcing}
If no external labels are admissible, then all qualitative structure is
quotient-invariant structure of the witness algebra.
\begin{proof}[Proof]
From $\noth$, there are no external labels, names, or tags.  A0$^*$ says a
distinction is real iff a finite endogenous replayable witness separates
it.  Two candidate objects $x, y$ that no admissible witness separates are
\emph{indistinguishable}:
\[
  x \sim y \;\;\iff\;\; \forall\, w \in \WitAlg,\;\;
  \mathrm{out}(w, x) = \mathrm{out}(w, y).
\]
Any property $P$ that assigns different values to indistinguishable objects
would require an external label to distinguish them---but no external
labels exist (from $\noth$).  By A0$^*$, the distinction ``$P(x) \neq P(y)$
despite $x \sim y$'' has no witness, hence is not real.  Therefore every
meaningful property is quotient-invariant:
\[
  x \sim y \;\;\Longrightarrow\;\; P(x) = P(y).
\]
\forcing{No external labels $\Rightarrow$ only witness-stable distinctions
$\Rightarrow$ qualitative quotient structure.  This forces: identity
(persistence under witnessing), objecthood (stable distinction class),
property (invariant of a class), relation (witnessable comparison class),
topology (observable-open structure from tests), causality (refinement
order in the closure boundary), modality (survivor-class structure of
continuations).}
\end{proof}
\end{theorem}

% ---------------------------------------------------------------------------
\subsection{Quantitative Forcing}
\label{sec:forcing:quantitative}

\begin{theorem}[Q2 --- Quantitative Forcing]
\label{thm:quantitative-forcing}
If no external units are admissible, then all quantitative structure is
determined only up to internal scale gauge, and only gauge-invariant
combinations are physically real.
\begin{proof}[Proof]
From $\noth$, there are no external units, absolute scales, or externally
chosen coordinates.  Any numerical assignment $f(x)$ that changes under
unit rescaling $d \to \lambda d$ is a bare number with no witnessable
content---the distinction ``$f = 3$ in units $u_1$'' vs.\ ``$f = 7$ in
units $u_2$'' has no witness unless a unit-independent comparison is
available.  By A0$^*$, only witnessable distinctions are real.  Therefore
only gauge-invariant combinations of quantities are physically meaningful.
\forcing{No external units $\Rightarrow$ only internally comparable
costs/frequencies/refinements $\Rightarrow$ quantitative invariants only
up to scale gauge.}
\end{proof}
\end{theorem}

\begin{theorem}[Q3 --- Gauge-of-Units]
\label{thm:gauge-of-units}
Any two quantitative presentations related by unit rescaling or coordinate
relabeling represent the same physical content iff all invariant ratios
and normalized evaluations agree.
\begin{proof}[Proof sketch]
This is the quantitative analogue of Q1.  Just as two qualitatively
indistinguishable objects are the same truth class, two quantitatively
gauge-related assignments are the same physical quantity.  The proof is
the same: distinguishing them would require an external unit standard,
which violates $\noth$.
\end{proof}
\end{theorem}

\begin{theorem}[Q4 --- Canonical Quantities]
\label{thm:canonical-quantities}
The only physically meaningful numbers are those invariant under the unit
gauge group.  Every admissible quantity is one of:
\begin{enumerate}[nosep, label=(\arabic*)]
  \item a dimensionless invariant,
  \item a ratio of like quantities,
  \item a generator of an internal comparison group,
  \item a normal form under scale gauge.
\end{enumerate}
\begin{proof}[Proof sketch]
Any quantity not of these forms depends on an external unit choice, which
is not witnessable from within.  By A0$^*$, it is not real.  This forces
every downstream quantity to be defined as an internal invariant:
\begin{itemize}[nosep]
  \item time from irreversible ledger accumulation (not external seconds),
  \item entropy from unresolved multiplicity (not external thermodynamics),
  \item energy from internal witness cost of stabilizing distinctions,
  \item distance from minimal separative comparison cost,
  \item probability from normalized positive state evaluation,
  \item curvature from variation of the internal distinguishability metric.
\end{itemize}
\forcing{No external scale $\Rightarrow$ every quantity is defined only up
to internal gauge, and only gauge-invariant combinations are real.}
\end{proof}
\end{theorem}

\begin{remark}[Bridge to Downstream Structure]
\label{rem:forcing-bridge}
Q1--Q4 are the bridge between nothingness and the derivation chain.  They
ensure that every structure derived in \cref{sec:derivation} and beyond is
not merely ``consistent with A0$^*$'' but \emph{forced} by it:
\begin{itemize}[nosep]
  \item The truth quotient (Step~6) is forced by Q1: indistinguishable
        objects must be identified.
  \item Gauge invariance (Step~12) is forced by Q1 + Q3: untestable
        relabelings are qualitative gauge.
  \item The witness-path metric (Step~16) is forced by Q4: distance is
        the minimal internal separative comparison cost.
  \item The conductance exponent $w = C/d^2$ (Step~28) is forced by Q2:
        scale covariance is a consequence of no external units.
  \item The Born rule (\cref{thm:born-rule}) is forced by Q4: probability
        is the unique normalized positive state evaluation.
\end{itemize}
\end{remark}

% ---------------------------------------------------------------------------
\subsection{The Static Universe}
\label{sec:forcing:static}

The $\Pit$-fixed closure whole is atemporal.  Time is a local phenomenon
arising from incomplete self-resolution, not a property of the universe
as a whole.

\begin{theorem}[S0 --- Whole-Before-Time]
\label{thm:whole-before-time}
The full $\Pit$-fixed closure whole $U = \mathrm{Fix}(\Pit)$ is
atemporal: applying $\Pit$ to $U$ changes nothing.  All admissible
local states are partial sections of~$U$.
\begin{proof}
By definition, $U = \mathrm{Fix}(\Pit)$ satisfies $\Pit(U) = U$.
Therefore no witness applied to $U$ can refine the truth quotient
further---the fiber $\fiber{p}$ is a singleton for every truth class
$p$ in~$U$, so no test produces new information.  By Step~8
(\cref{thm:time}), time is the irreversible formation of new records,
measured by $\Del\Tim = \log(|\fiber{}|_{\mathrm{pre}} /
|\fiber{}|_{\mathrm{post}}) \geq 0$.  When the fiber is already
minimal ($|\fiber{}| = 1$), every test yields $\Del\Tim = 0$: no time
passes.  The whole $U$ is therefore atemporal.

Any admissible local state $s$ satisfies $s \subseteq U$ (since $U$
contains all admissible truth classes) but in general
$\Pit(s) \neq s$---the local truth quotient is not yet fully resolved.
Hence $s$ is a partial section of the completed whole.
\forcing{$U = \mathrm{Fix}(\Pit)$ is forced by the closure
construction.  Atemporality follows from Step~8: when nothing remains
to resolve, $\Del\Tim = 0$ identically.  Local states are partial
sections because they are subsets of the complete truth quotient that
have not yet exhausted their witness budget.}
\end{proof}
\end{theorem}

\begin{theorem}[S1 --- Time is Local Defect-Resolution]
\label{thm:time-local-defect}
If $s$ is a local incomplete section of $U$, then ``time'' is the
monotone reduction of unresolved witness-defect
$\ClosureDefect{Q}(s) = \Pit(s) \setminus s$, not motion of $U$
itself.  The universe is static globally, dynamic only as local
self-resolution.
\begin{proof}
By \cref{thm:closure-defect} (Step~23), the closure-defect
$\ClosureDefect{Q}(s) = \Pit_Q(s) - s$ measures the gap between the
current state and its truth projection.  Its support
$\Residue_t = \mathrm{supp}(\ClosureDefect{Q}(s_t))$ decreases
monotonically with each test (Step~8): $|\Residue_{t+1}| \leq
|\Residue_t|$.  This monotone decrease \emph{is} time for the local
section~$s$.

For the global whole $U = \mathrm{Fix}(\Pit)$, the closure-defect
vanishes: $\ClosureDefect{Q}(U) = \Pit(U) - U = 0$.  Therefore $U$
has no unresolved components and no time evolution.  ``Dynamics'' is
exclusively a local phenomenon: the self-resolution of a partial
section's witness-defect toward the already-existing fixed whole.
\forcing{The global whole is static because $\ClosureDefect{Q}(U) = 0$
(no defect to resolve).  Local time is the monotone decrease of
$|\Residue_t|$, which is forced by the append-only ledger (Step~5)
and the irreversibility of record formation (Step~8).  The universe
does not ``move''---local sections resolve toward it.}
\end{proof}
\end{theorem}

% ---------------------------------------------------------------------------
\subsection{The Quantitative Seed}
\label{sec:forcing:seed}

The structural forcing of Sections~5.1--5.3 determines the \emph{form} of all
physical law.  To pin down the \emph{values} of dimensionless observables, we
isolate the unique quantitative seed $\delta^*$.

\begin{definition}[Defect Space and Action]
Let $D$ denote the space of closure-defect profiles (lists of
component-state fibers).  The gauge group $G$ acts by permuting
distinction labels, preserving total defect.  The \emph{action
functional} $A : D/{\sim_G} \to \mathbb{N}$ maps each gauge class to
its total defect.  A defect is \emph{self-retaining} if every open
component has fiber~$1$ (minimum nonzero), making it a fixed point
under witness accumulation.  A defect is \emph{non-gauge} if $A(d) > 0$.
\end{definition}

\begin{theorem}[Seed Existence and Uniqueness]
\label{thm:seed-existence}
There exists a unique (up to gauge) self-retaining non-gauge defect
$\delta^*$ minimizing $A$.  Concretely: $A(\delta^*) = 1$, and every
non-gauge defect has $A \ge 1$.
\begin{proof}
The class $D_\text{sr}$ of self-retaining non-gauge defects is nonempty
(witness: a single open component with fiber~$1$).  The action maps to
$\mathbb{N}$, which is well-ordered.  A minimizer exists by
well-foundedness on a nonempty set.  Any two minimizers have the same
action value by symmetric minimality.
\end{proof}
\end{theorem}

\begin{theorem}[Seed is Renormalization Fixed Point]
\label{thm:seed-fixed}
For any refinement operator $R$ that is monotone, gauge-preserving,
and self-retaining-preserving, $A(R(\delta^*)) = A(\delta^*)$.
\begin{proof}
By monotonicity $A(R(\delta^*)) \le A(\delta^*)$; by minimality
$A(\delta^*) \le A(R(\delta^*))$.
\end{proof}
\end{theorem}

% ---------------------------------------------------------------------------
\subsection{Quantitative Closure}
\label{sec:forcing:closure}

The linearized refinement operator $L^* = DR|_{\delta^*}$ acts on the
tangent space at the seed.  Its spectral, holonomy, and normal-form data
determine every dimensionless observable.

\begin{theorem}[Dimensionless Observable Decomposition]
\label{thm:observable-decomp}
Every dimensionless observable $\mathcal{O}$ satisfies
\[
  \mathcal{O} = F\bigl(\operatorname{Spec}(L^*),\;
  \operatorname{Hol}(L^*),\;\operatorname{NF}(L^*)\bigr).
\]
\begin{proof}
Six-step ladder:
\emph{(1)}~gauge invariance $\Rightarrow$ quotient-defined;
\emph{(2)}~quotient $\Rightarrow$ seed-determined (via uniqueness);
\emph{(3)}~linear invariants $=$ spectral data;
\emph{(4)}~gauge-sector invariants $=$ holonomy;
\emph{(5)}~interaction invariants $=$ normal-form coefficients;
\emph{(6)}~corollary: $\mathcal{O}$ factors through
$(\operatorname{Spec}, \operatorname{Hol}, \operatorname{NF})$.
\end{proof}
\end{theorem}

Dimensional quantities (mass, coupling, $\Lambda$) are fixed up to the
internally derived normalization map $(\hbar^* = A(\delta^*),\;
c^* = \sup d_\text{sep}/\Delta T)$.  No external unit system enters.
