% =============================================================================
% Section 1: Introduction
% =============================================================================

\section{Introduction}
\label{sec:introduction}

\subsection{The Question}

What is the minimal structure forced into existence by the requirement that
every distinction be witnessable?

This is not a question about engineering, artificial intelligence, or any
particular domain of application. It is a foundational question about
describable reality. If we insist, as A0$^*$ (Completed Witnessability) requires, that every meaningful
distinction be backed by a finite procedure that can in fact separate it, then
some structures become unavoidable. The purpose of this paper is to identify
those structures and to make explicit which commitments are definitional and
which are derived.  Crucially, A0$^*$ is itself \emph{derived} from
nothingness ($\noth$) through a chain of five necessity theorems
(N0--N4)---the paper assumes nothing and derives its operating principle.

The central result is a 34-step derivation that begins at $\noth$ and
unfolds by reapplying A0$^*$ to its own products, tracing a single reflective
chain: nothingness forces closure, closure reveals defect, defect requires
self-observation, and self-observation iterated on the continuum limit
produces geometry.  Within the paper's model class, every structure in the
chain---from the carrier of descriptions to the K\"{a}hler triple on the
truth-quotient limit---is forced up to computable isomorphism; none is
introduced as a free design parameter.

Equally important is what we do \emph{not} claim. We do not claim to derive the
specific \emph{content} of physical law, only a vocabulary and scaffolding for
reasoning under witnessability. We do not claim that every modeling choice is
eliminated under every possible formalization of A0$^*$. Our claim is narrower and
testable: given the explicit premises and conventions in \cref{sec:axioms} and
the model class studied in \cref{sec:derivation}, the resulting chain of
structures is canonical up to computable isomorphism.

\subsection{The Witnessability Principle}

We begin from a single principle, derived from $\noth$ (not postulated):

\begin{quote}
\textbf{A0 (Bare Witnessability).}  A distinction is admissible if and
only if a finite witness procedure can separate it.
\end{quote}

\noindent
The completed form, A0$^*$, adds a reflective clause: witness validity itself must be witnessable by a process satisfying the same criterion.  A0$^*$ is not an axiom but the first theorem of the framework, derived from $\noth$ via five necessity theorems (N0--N4, see \cref{sec:axioms:derivation}).  All subsequent references use A0$^*$.

This principle is not chosen for convenience. It is the weakest non-trivial
constraint one can impose. If a distinction cannot even in principle be
witnessed, then it is operationally meaningless. Wheeler's ``It from Bit''
programme \citep{wheeler1990} and Rovelli's relational quantum
mechanics \citep{rovelli1996} both point toward observer-relative,
information-theoretic foundations. A0$^*$ isolates the operational core shared by
such views and elevates it to an admissibility criterion.

The formal statement of A0$^*$ and its immediate consequences are given in
\cref{sec:axioms}.

\subsection{Key Insight: Nothingness as the Only Unfalsifiable Truth}

Nothingness, the state carrying zero information, asserting nothing, and
distinguishing nothing, is the only unfalsifiable truth. Every other claim can
in principle be tested and potentially refuted. Nothingness alone survives all
tests trivially because it makes no assertions.

This claim is meant operationally. $\noth$ is the unique state satisfying
$\Del = \varnothing$, so no tests exist, no descriptions exist, and no
distinctions are drawn. From this state, the self-application of A0$^*$ forces the
existence of at least one admissible distinction, which breaks the symmetry.
Once at least one distinction exists, the derivation proceeds step by step,
making explicit where results are unique up to computable isomorphism and
where the development depends on declared conventions.

\begin{enumerate}[label=\textbf{S\arabic*}.]
  \item $\noth$ is the unique unfalsifiable state; self-application of A0$^*$
        forces the existence of at least one admissible distinction, breaking
        the symmetry.
  \item A0$^*$ applied to its own products yields, step by step, the three
        primitives ($\Del$, $\Pit$, $\Tim$), a carrier closed under
        witnessing \citep{church1936, turing1936}, gauge invariance,
        separator geometry, Bellman-optimal dynamics, self-hosting closure,
        and a K\"{a}hler structure on the continuum limit.
\end{enumerate}

The forcing claim is therefore precise. We claim canonical structure within
the explicitly stated model class and up to computable isomorphism, and we
state frontiers where additional assumptions, alternative modeling choices, or
empirical input would be required.  \Cref{fig:phase-overview} provides a
roadmap of the nine phases.

\input{figures/phase-overview}

\subsection{Contributions}

This paper makes three contributions:

\begin{enumerate}[label=(\arabic*)]
  \item \textbf{A 34-step derivation of structural reality from $\noth$.}
        Starting from nothingness, we derive the framework's core
        vocabulary across nine phases (A--I), introducing 25 new
        mathematical objects.  No external
        assumption beyond A0$^*$ enters the derivation; the Church--Turing
        thesis independently corroborates Step~3's mathematical conclusion.
        A three-point self-audit and Z3 verification of selected
        algebraic properties on finite models accompany the derivation
        (\cref{sec:derivation}, Appendix~\ref{app:derivation},
        Appendix~\ref{app:z3}).

  \item \textbf{Three forced primitives and seven doctrines.}
        The derivation isolates three primitives, $\Del$ (tests),
        $\Pit$ (truth), and $\Tim$ (time), and collects seven doctrines that
        follow from the operational reading of A0$^*$
        (\cref{sec:primitives}, \cref{sec:doctrines}).

  \item \textbf{K\"{a}hler geometry, Born rule, and the split law.}
        The continuum limit of truth quotients carries a forced
        K\"{a}hler triple $(g, J, \omega)$, a unique gauge-invariant
        probability (the Born rule of the framework), and a canonical
        dynamical decomposition whose sector-readouts recover the forms
        of quantum mechanics, gauge theory, and thermodynamics
        (\cref{sec:derivation}).
\end{enumerate}

\subsection{Paper Outline}

\Cref{sec:axioms} derives A0$^*$ from $\noth$ and states the forced primitives.
\Cref{sec:primitives} introduces the three forced primitives ($\Del$,
$\Pit$, $\Tim$).
\Cref{sec:doctrines} presents the seven non-negotiable doctrines.
\Cref{sec:derivation} presents the 34-step derivation (Phases A--I) with proof sketches.
\Cref{sec:verification} reports the three-point self-audit and Z3 verification.
\Cref{sec:related} surveys the intellectual context.
\Cref{sec:discussion} provides an honest assessment of claims, limitations,
and open frontiers.
\Cref{sec:conclusion} concludes with future directions.
Appendices contain the complete derivation with full proofs, verification
protocol, and Z3 scripts.
