% =============================================================================
% Section: The Three Primitives
% =============================================================================

\section{The Three Primitives}
\label{sec:primitives}

The derivation from nothingness forces exactly three primitive concepts into
existence.  They are not postulated; they are the unique consequences of
applying A0$^*$ to its own products.  This section names them, states their
definitions, and traces each to the derivation step where it first appears.

% -------------------------------------------------------------------------
\subsection{$\Del$ --- Tests (Distinctions)}
\label{sec:primitives:delta}

\begin{definition}[Tests]
\label{def:prim-delta}
A \emph{test} $\tau \in \Del$ is a total computable function that separates
descriptions.  The set of all tests $\Del^*$ is the least fixed point of
closure rules on the base tests $\Del_0$, enumerated canonically by
$\CanEnum$ (Step~4, \cref{thm:endogenous}).
\end{definition}

Tests are the only epistemic currency of the framework.  A distinction
without a separating test is operationally void---it violates A0$^*$.
Every claim, every piece of evidence, every advance in knowledge reduces to
the execution of a test and the recording of its outcome.  Tests emerge at
Steps~3--4 of the derivation: Step~3 forces the computational substrate
that can execute them, and Step~4 forces their closure into a self-generating
algebra $\Del^*$.

% -------------------------------------------------------------------------
\subsection{$\Pit$ --- Truth}
\label{sec:primitives:pi}

\begin{definition}[Truth]
\label{def:prim-pi}
\emph{Truth} is the information that survives all valid tests.  Formally,
the \emph{truth quotient} $\TQ{\Ledger} = \FinSlice / {\equivL}$ identifies
descriptions that no recorded test can distinguish (Step~6,
\cref{thm:truth-quotient}).
\end{definition}

Truth is not a Platonic ideal; it is the coarsest structure that survives
all executed tests.  Two descriptions are ``the same truth'' precisely when
no test separates them.  The truth quotient carries additional forced
structure: an observable-open topology $\ObsOpen$, an eraser algebra
$\Eraser$, and a self-generating fixed point $\mathrm{sim}^*$ (Step~7,
\cref{thm:indistinguishability}).  Together these make truth not merely a
quotient but a topological algebra of testability.
The concrete algorithm that computes canonical representatives---selecting the lexicographically minimal serialization within each truth class---is $\PiCanon$-canonicalization (Definition~\ref{def:pi-canon}).

% -------------------------------------------------------------------------
\subsection{$\Tim$ --- Time}
\label{sec:primitives:time}

\begin{definition}[Time]
\label{def:prim-time}
\emph{Time} is the irreversible formation of records.  Each new entry
appended to the ledger $\Ledger$ advances the clock by
$\Del\Tim = \log(|\fiber{}|_{\mathrm{pre}} / |\fiber{}|_{\mathrm{post}})
\geq 0$ (Step~8, \cref{thm:time}).
\end{definition}

Time is not a parameter imposed from outside; it \emph{is} the
irreversibility of witnessing.  The ledger is append-only (Step~5): deleting
a record would destroy a witness, violating A0$^*$.  Because each record can
only refine the truth quotient (never coarsen it), the sequence of records
defines an arrow---from less resolved to more resolved---that cannot be
reversed.  This arrow is time.  The energy ledger $\EnergyLedger$ (Step~10)
tracks the cost of each advance.

% -------------------------------------------------------------------------
\subsection{$\Truthpoint$ --- The Truthpoint}
\label{sec:primitives:truthpoint}

\begin{definition}[Truthpoint]
\label{def:truthpoint}
The \emph{truthpoint} $\Truthpoint$ is the limit of $\TQ{\Ledger}$ as the
ledger grows without bound:
\[
  \Truthpoint \;=\; \lim_{|\Ledger| \to \infty} \TQ{\Ledger}.
\]
It is the ideal complete knowledge state---the truth quotient under
unlimited budget---approached but never operationally reached.
\end{definition}

$\Truthpoint$ exists as a limit concept (Step~11, \cref{thm:opnoth}): when
the budget is exhausted, the system reaches operational nothingness $\opnoth$,
the horizon of resolvability.  The gap between any finite $\TQ{\Ledger}$ and
$\Truthpoint$ is precisely what the $\OMEGA$ state reports: the surviving
ambiguity, together with the minimal separator test that would reduce it.

\bigskip
\noindent
\textbf{Summary.}  These three primitives---$\Del$ (tests), $\Pit$ (truth),
$\Tim$ (time)---together with the limit concept $\Truthpoint$, are not design
choices.  They are the unique forced consequences of A0$^*$ applied to its own
products.  Every subsequent structure in the derivation (gauge invariance,
Myhill--Nerode congruence, separator geometry, Bellman dynamics,
self-hosting) is built from these three primitives and nothing else.
