\section{Z3 Formal Verification}
\label{app:z3}

This appendix documents the SMT-LIB verification scripts used to check
the algebraic properties of the derivation on finite models.  All 20
scripts were executed using Z3 \citep{demoura2008} v4.15.0 and are
available in the \texttt{scripts/z3/} directory.

\subsection{Scope and Limitations}

\textbf{What is verified:} Structural properties that can be expressed as
first-order constraints over finite domains---ledger append-only semantics,
equivalence relation axioms, eraser algebra, monotone fiber shrinkage,
gauge group structure ($\mathrm{Sym}(2) \times \mathrm{Sym}(3)$ with full
Cayley table), image factorization, right-congruence (including coarsest
congruence), separator-induced metric (with (TS)-necessity demonstration),
Bellman well-definedness with non-degenerate Coupling Law, fixed-point
consistency, trichotomy exhaustiveness, and dependency DAG acyclicity.

\textbf{What is NOT verified:} Infinite-domain properties (e.g., the full
topological structure of $\ObsOpen$, the real-analytic properties of
$\SepDist$ on continuous spaces), the naming convention for the forced class, and semantic forcing arguments (``this is the \emph{unique}
A0-compliant continuation'').  These require mechanization in a proof
assistant (Lean, Coq) or human peer review.

\textbf{Honest disclosure:} Z3 checks that the stated properties are
\emph{satisfiable} on finite models (i.e., a model exists).  This is
weaker than proving the properties hold \emph{for all} models.  Where
we check unsatisfiability (e.g., ``there is no model violating the
triangle inequality''), we state this explicitly.

\textbf{Clarification on scope:} These scripts are best understood as
\emph{machine-checked sanity checks on finite instances}, not as
general proofs of the universal theorems stated in the derivation.  Each
script operates on a small domain (typically 3--8 elements) and verifies
that the claimed algebraic structure is realizable.  Proving the general
theorems---that the properties hold for \emph{all} finite models of the
appropriate size, and that the forcing arguments produce \emph{unique}
A0-compliant continuations---requires mechanization in a full proof
assistant (Lean, Coq, or Isabelle), which is deferred to future work.

\subsection{Scripts and Results}

\begin{table}[h!]
\centering
\small
\begin{tabular}{@{}lclp{5.5cm}@{}}
\toprule
\textbf{Script} & \textbf{Step} & \textbf{Parts} & \textbf{Property Verified} \\
\midrule
\texttt{step05-append-only} & 5 & 3 & Append-only ledger; removal destroys witnesses (violates A0); Diamond Law \\
\texttt{step06-equiv} & 6 & 3 & Equivalence relation (R, S, T) on 4 elements; coarsest quotient \\
\texttt{step07-eraser} & 7 & 3 & Observable-open intersection closure; eraser idempotence; eraser commutativity \\
\texttt{step08-monotone} & 8 & 3 & Fiber shrinkage: $|W'| \leq |W|$ \\
\texttt{step12-groupoid} & 12 & 5 & $\mathrm{Sym}(2) \times \mathrm{Sym}(3)$ group: 12 elements, full Cayley table, closure, inverses, associativity, no 13th element \\
\texttt{step14-split} & 14 & 3 & Image factorization $f = \iota \circ q$ \\
\texttt{step15-myhill-nerode} & 15 & 4 & Right-congruence on 3-state DFA; coarsest congruence (merging distinguishable states is unsat) \\
\texttt{step16-metric} & 16 & 4 & Separator-induced metric: min-cost separators satisfy triangle inequality; (TS)-violating costs \emph{can} break it \\
\texttt{step18-bellman} & 18 & 7 & Bellman recursion with \emph{different} gains/costs; argmax efficiency; $\Pit$-canonical tie-break; policy invariance; determinism \\
\texttt{step19-fixpoint} & 19 & 3 & $\FixPt(\mathcal{S}) = \mathcal{S}$ \\
\texttt{step21-dichotomy} & 21 & 8 & Trichotomy exhaustive \\
\texttt{step22-inverse-limit} & 22 & 3 & Metric preservation in inverse system \\
\texttt{step23-invariant-measure} & 23 & 3 & Uniqueness of Sym-invariant counting measure \\
\texttt{step24-dirichlet} & 24 & 4 & Dirichlet form axioms (E1)--(E3) on finite graph \\
\texttt{step25-semigroup} & 25 & 4 & Markov semigroup properties \\
\texttt{step26-intrinsic-metric} & 26 & 4 & Intrinsic distance metric axioms \\
\texttt{step27-complex-structure} & 27 & 3 & $J^2 = -\mathrm{id}$ on 4D model \\
\texttt{step28-symplectic} & 28 & 3 & Closedness on simplicial complex \\
\texttt{step29-kahler} & 29 & 3 & K\"ahler compatibility conditions \\
\texttt{step-dag-acyclic} & --- & 3 & Dependency DAG (34 steps): topological order exists; back-edge creates cycle; all steps reachable \\
\bottomrule
\end{tabular}
\caption{Z3 verification results (20 scripts, 77 total check-sat calls).  Each script verifies multiple parts; all expected sat/unsat results match.}
\label{tab:z3-results}
\end{table}

\subsection{Running the Verification}

\begin{verbatim}
  ./scripts/z3/run-all.sh
\end{verbatim}
Expected output: 20/20 pass.

\subsection{Key Strengthening (v2)}

Three proofs were substantially strengthened from the initial version:

\begin{itemize}
  \item \textbf{Step~18 (Coupling Law):} The original proof used equal gains
    for all tests, reducing \texttt{argmax}($\varepsilon$) to
    \texttt{argmin}(cost)---a degenerate case.  The new version uses 4 tests
    with different gains and costs, verifies $\Pit$-canonical tie-breaking,
    policy invariance under strictly monotone transformations, and uniqueness
    of selection.

  \item \textbf{Step~12 (Gauge Group):} The original checked group axioms on
    3 abstract morphisms.  The new version explicitly constructs
    $\mathrm{Sym}(2) \times \mathrm{Sym}(3)$ with all $2! \times 3! = 12$
    elements, a complete $6 \times 6$ Cayley table for $S_3$, and proves no
    13th element can exist.

  \item \textbf{Step~16 (Separator Metric):} The original checked generic
    metric axioms on abstract points.  The new version models 3 truth classes
    with 3 tests of known costs, computes $d(p,p') = \min\{c(\tau) : \tau
    \text{ separates } p, p'\}$, and demonstrates that (TS)-violating cost
    models \emph{can} break the triangle inequality (showing (TS) is
    necessary, not just sufficient).
\end{itemize}

Three new proofs were added:

\begin{itemize}
  \item \textbf{Step~5 (Append-Only Ledger):} Verifies that removing a
    ledger entry destroys a witness, violating A0.  Also checks Diamond Law
    (order-independence).

  \item \textbf{Step~7 (Eraser Algebra):} Verifies observable-open
    intersection closure, eraser idempotence, and eraser commutativity on 4
    truth classes with 3 tests.

  \item \textbf{DAG Acyclicity:} Encodes all 34 steps with their declared
    dependencies, verifies a topological ordering exists, adding a back-edge
    creates a cycle, and all steps are downstream of Step~0.
\end{itemize}

\subsection{Phase H Proofs (v3)}

Eight new scripts verify the geometric reconstruction (Steps~22--29).
These proofs operate on finite discrete models and check the algebraic
properties that underpin the continuum-limit argument:

\begin{itemize}
  \item \textbf{Step~22 (Inverse Limit):} Verifies metric preservation
    in a finite inverse system of truth-quotient graphs, confirming that
    refinement maps are isometric embeddings.

  \item \textbf{Step~23 (Invariant Measure):} Checks uniqueness of the
    Sym-invariant counting measure on a 4-element set, verifying that
    the uniform measure is the only gauge-invariant probability measure.

  \item \textbf{Step~24 (Dirichlet Form):} Verifies the three Dirichlet
    form axioms (E1: bilinearity, E2: Markov property, E3: closedness)
    on a weighted graph with conductance $w = 1/d^2$.

  \item \textbf{Step~25 (Semigroup):} Checks Markov semigroup properties
    (positivity preservation, contraction, composition law) derived from
    the Beurling--Deny theorem applied to the Dirichlet form.

  \item \textbf{Step~26 (Intrinsic Metric):} Verifies that the intrinsic
    distance induced by the Dirichlet form satisfies metric axioms
    (non-negativity, symmetry, triangle inequality, separation).

  \item \textbf{Step~27 (Complex Structure):} Checks $J^2 = -\mathrm{id}$
    and the Nijenhuis integrability condition on a 4-dimensional model,
    confirming that the almost-complex structure is integrable.

  \item \textbf{Step~28 (Symplectic Form):} Verifies closedness and
    non-degeneracy of the symplectic 2-form on a simplicial complex,
    using the discrete exterior derivative.

  \item \textbf{Step~29 (K\"ahler):} Checks the K\"ahler compatibility
    conditions $g(JX, JY) = g(X, Y)$ and $\omega(X, Y) = g(JX, Y)$,
    confirming that the metric, complex structure, and symplectic form
    are mutually compatible.
\end{itemize}

\subsection{Verification Boundary}

\begin{table}[h!]
\centering
\small
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{Can verify (Z3, finite models)} & \textbf{Cannot verify (needs Lean/Coq)} \\
\midrule
Equivalence relation axioms & Forcing (``unique continuation'') \\
Group structure ($\prod \mathrm{Sym}$) & Infinite-domain theorems \\
Monotone fiber shrinkage & Observable-open topology \\
Image factorization & Separator geometry (real analysis) \\
Myhill--Nerode congruence & Semantic ``why this is forced'' \\
Separator metric (finite) & \\
Bellman + Coupling Law & \\
Eraser algebra & \\
Ledger append-only & \\
DAG acyclicity (34 steps) & \\
Fixed-point consistency & \\
Trichotomy exhaustiveness & \\
Dirichlet form axioms & Continuum limit (inverse system) \\
Riemannian metric (intrinsic) & Smooth manifold structure \\
Complex structure ($J^2=-\mathrm{id}$) & K\"ahler compatibility \\
\bottomrule
\end{tabular}
\caption{Boundary between finite-model Z3 verification and infinite-domain mechanization.}
\label{tab:z3-boundary}
\end{table}
